### 2022-07-19 manuscript/manuscript.Rmd
update ms
- add methods that were missing

**Authors**: Patricio Barría,^1^ Matías Castillo-Aguilar,^2,3^ Cristian Núñez-Espinoza,^3,4^ \*

@@

Nevertheless, and to the best of our knowledge, there is no robust evidence characterising the observed variation in developmental traits in infants with known CH across age. Therefore, our main objective in this study was to describe and model the relationship between sociodemographics, prematurity and neurodevelopmental levels based on ASQ-3 scores in infants with diagnosed CH.

@@

[¿Los datos vienen de algún grupo de estudio, cohorte o de algún proyecto en particular? Detalles de donde vino la muestra]{custom-style="ErrorTok"}

[¿Qué resguardos éticos se consideran? Comité de ética y otros permisos que sean pertinentes]{custom-style="ErrorTok"}

@@

```{r}
# Sample
n_sample <- dataset[, .N]
#> 234

# Sample by sex
n_sex <- dataset[j = .(stat = paste0(.N, " (", round(.N*100/n_sample, 1), "%)")), 
                 keyby = .(sex = sexo_paciente)]
#>    sex        stat
#> 1:   F  94 (40.2%)
#> 2:   M 140 (59.8%)
```

[¿Cómo, cuando y bajo qué condiciones fueron reclutados los sujetos? Criterios de elegibilidad, tipo de muestreo, etc.]{custom-style="ErrorTok"}

[¿Cómo, cuando y quien hizo el diagnóstico de hipotonía congénita? ¿Fue igual para todos? ¿La información vino de parte de terceros?]{custom-style="ErrorTok"}

@@

The ASQ-3 is a parent reported initial level developmental screening instrument consisting of 21 intervals, each with 30 items in five areas: i) communication (CM), ii) gross motor (GM), iii) fine motor (FM), iv) problem-solving (CG), and v) personal-social (PS) [@squires2009ages]. The ASQ is cost-effective and widely used in the United States and other countries [@sarmiento2011universal; @heo2008cross]. It has been translated into several languages, and the number of international studies on its psychometric properties with diverse cultural environments is increasing [@sarmiento2011universal; @heo2008cross]. It has shown adequate psychometric properties (75% sensitivity and 81% specificity) in Chilean term and preterm infants [@schonhaut2013validity].

@@

@@

[¿Como, cuando y quién recolectó los datos demográficos?]{custom-style="ErrorTok"}

@@

[¿Cómo se evaluó el ASQ-3? ¿Con qué protocolo? ¿Que consideraciones se tuvieron en la evaluación? ¿Algo más que podría ser relevante para agregar respecto a la evaluación o corrección del cuestionario?]{custom-style="ErrorTok"}

@@

```{r}
# Prop test on sex
prop_sex <- dataset[, table(sexo_paciente)] |> 
  chisq.test() |> 
  report_chi(effsize = FALSE)
#> [1] "$\\chi^2_{gof}$ (1) = 9.04, *p* = 0.003, $C_{Pearson}$ = 0.19, CI~0.95%~[0.09, 1]"
```

@@

>>>>

**Authors**: Patricio Barría^1^, Matías Castillo-Aguilar^2,3^, Katherine Harris^2,3^, Cristian Núñez-Espinosa^3,4^ \*

@@

Nevertheless, and to the best of our knowledge, there is no robust evidence characterising the observed variation in developmental traits in infants with diagnosed CH across age. Therefore, our main objective in this study was to describe and model the relationship between sociodemographics, prematurity and neurodevelopmental levels based on ASQ-3 scores in infants with diagnosed CH.

@@

@@

```{r}
# Sample
n_sample <- dataset[, .N]
#> 234
```

A total of `r n_sample` patients were enrolled as part of an intervention programme carried out by the *Cruz del Sur* Rehabilitation Centre (Punta Arenas, Chile), admitted from one month of age to 60 months, and assessed at admission, control and discharge. The total records of patients diagnosed with CH admitted to the institution's programme were analysed and the assessment of the subjects was conducted and guided by a nurse trained in the application of the ASQ-3.

In this context, it is worth noting that the data, as well as the patient registry, are part of institutional strategies aimed at the diagnosis and continuous monitoring of the clinical situation of users, which are used to improve care processes and clinical decision-making.

The diagnosis of CH is based on 4 sources: 1) primary care paediatrician, who refers the diagnosis to the institution, where admission to the programme is made; 2) paediatricians from private clinics, who refer the diagnosis; 3) neuropaediatricians from the clinical hospital, who refer the diagnosis to the programme; 4) physiatrist from the institution, who assigns the diagnosis if it corresponds to the semiology.

@@

The ASQ-3 is a parent-reported initial level developmental screening instrument, consisting of 21 intervals, each with 30 items in five areas: i) communication (CM), ii) gross motor (GM), iii) fine motor (FM), iv) problem-solving (CG), and v) personal-social (PS) [@squires2009ages]. The ASQ is cost-effective and widely used in the United States and other countries [@sarmiento2011universal; @heo2008cross]. It has been translated into several languages, and the number of international studies on its psychometric properties with diverse cultural environments is increasing [@sarmiento2011universal; @heo2008cross]. It has shown adequate psychometric properties (75% sensitivity and 81% specificity) in Chilean term and preterm infants [@schonhaut2013validity].

@@

For the collection of research data, an authorisation consent was obtained for the use of instrumental clinical data as well as the clinical record of each patient. Subsequently, and in accordance with national research regulations, written authorisation was obtained from the institutional director for the use of the records and database for research purposes. To this end, the data were anonymised during data processing and subsequent analyses.

@@

The demographic data of each patient was collected and made available throughout the study by the institutional electronic systems at the time of entry to the programme. The administrative registration was done by the secretaries of the user coordination unit, recording name, ID number, date of birth and diagnosis. In addition, age data was automatically updated by the computer system's algorithms.

@@

The ASQ-3 was applied according to the protocol established by the instrument itself, with a face-to-face or telematic application being valid. The protocol can be found in the instrument's manuals, which were administered by the programme nurse. Some important considerations of the administration are that i) the main caregiver must respond, ii) in case of doubts of the caregiver about the assessed behaviour, the information is corroborated by an in situ test with the user, and iii) in relation to the correction, these procedures are standardised by the test itself.

@@

```{r}
# Sample by sex
n_sex <- dataset[j = .(stat = paste0(.N, " (", round(.N*100/n_sample, 1), "%)")), 
                 keyby = .(sex = sexo_paciente)]
#>    sex        stat
#> 1:   F  94 (40.2%)
#> 2:   M 140 (59.8%)

# Prop test on sex
prop_sex <- dataset[, table(sexo_paciente)] |> 
  chisq.test() |> 
  report_chi(effsize = FALSE)
#> [1] "$\\chi^2_{gof}$ (1) = 9.04, *p* = 0.003"
```

@@

No existe financiamiento extraordinario a la aplicación del instrumento.
